Designing Types for R, Empirically

### 2020-05-15 paper/main.tex
corpus table

\newcommand{\NUMFUNCTIONS}{25,000\xspace}  %%% TODO auto-gen
\newcommand{\NUMPACKAGES}{412\xspace}  %%% TODO auto-gen
\newcommand{\PACKAGES}{412\xspace}  %%% TODO auto-gen
\newcommand{\genthat}{genthat\xspace}  %%% TODO auto-gen
\newcommand{\YEARS}{20\xspace} %%TODO fix
\newcommand{\PERCFAILEDASSERTIONS}{2.19\%\xspace}
\newcommand{\PERCFAILEDASSERTIONSWUNDEF}{2.23\%\xspace}
\newcommand{\PERCASSERTIONSUNDEF}{0.04\%\xspace}
\newcommand{\NUMPKGSEVAL}{7485\xspace} % autogenerate
\newcommand{\NUMASSERTIONS}{62,171,573\xspace} %autogenerate
\newcommand{\PROPFUNSFAILEDCHECK}{21.21\%\xspace}
\newcommand{\PROPFUNSFAILEDCHECKNOSTHREE}{15.48\%\xspace} 
\newcommand{\PROPARGSFAILINGASSERTS}{3.53\%\xspace} % autogen....
\newcommand{\PERCSUCCARG}{96.46\%\xspace}
\newcommand{\PERCSUCCFUNNOSTHREE}{84.52\%\xspace}
\newcommand{\PERCSUCCFUNS}{78.79\%\xspace}
\newcommand{\PERCCALLSBADFUNSINTESTS}{5.38\%\xspace}
\newcommand{\TOTNUMSIGS}{21,730\xspace}

@@

>>>>

\newcommand{\NUMFUNCTIONS}{20,214\xspace}  %%% TODO auto-gen
\newcommand{\NUMPACKAGES}{412\xspace}  %%% TODO auto-gen
\newcommand{\PACKAGES}{412\xspace}  %%% TODO auto-gen
\newcommand{\genthat}{genthat\xspace}  %%% TODO auto-gen
\newcommand{\YEARS}{20\xspace} %%TODO fix
\newcommand{\PERCFAILEDASSERTIONS}{2.19\%\xspace}
\newcommand{\PERCFAILEDASSERTIONSWUNDEF}{2.23\%\xspace}
\newcommand{\PERCASSERTIONSUNDEF}{0.04\%\xspace}
\newcommand{\NUMPKGSEVAL}{7485\xspace} % autogenerate
\newcommand{\NUMASSERTIONS}{62,171,573\xspace} %autogenerate
\newcommand{\PROPFUNSFAILEDCHECK}{21.21\%\xspace}
\newcommand{\PROPFUNSFAILEDCHECKNOSTHREE}{15.48\%\xspace} 
\newcommand{\PROPARGSFAILINGASSERTS}{3.53\%\xspace} % autogen....
\newcommand{\PERCSUCCARG}{96.46\%\xspace}
\newcommand{\PERCSUCCFUNNOSTHREE}{84.52\%\xspace}
\newcommand{\PERCSUCCFUNS}{78.79\%\xspace}
\newcommand{\PERCCALLSBADFUNSINTESTS}{5.38\%\xspace}
\newcommand{\TOTNUMSIGS}{21,730\xspace}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Overview of Types for R}

We ran \typetracer on the \CorpusLoadable packages we discussed in this section, which yields types for \NUMFUNCTIONS functions.
To build up a picture of what types for R functions might look like, consult Table~\ref{tbl:corpus-type-breakdown}, which lists the 10 most popular argument signatures.
The first row of the table reads: 13255 arguments have type \dbl, represnting 11.06\% of type signatures, and \typetracer observed 41M values of type \dbl flowing into or out of functions while running on the core package corpus, which represents only 1.59\% of dynamically observed values.
Of interest are rows 2 and 4, where we see that collectively \lgl and \chr make up only 16.11\% of argument types, while making up nearly 60\% of dynamically observed values.

\begin{tabular}{lrrrr}
\toprule
Type & Args with Type & \% of Static Types & Dynamic Observations & \% of Dynamic Types\\
\midrule
double & 13,255 & 11.06 & 41,560,797 & 1.59\\
logical & 10,142 & 8.46 & 712,895,959 & 27.21\\
null & 9,175 & 7.66 & 122,202,183 & 4.66\\
character & 9,171 & 7.65 & 794,058,588 & 30.31\\
double[] & 7,811 & 6.52 & 29,272,232 & 1.12\\
\addlinespace
... & 7,759 & 6.48 & 76,369,333 & 2.92\\
any & 6,243 & 5.21 & 255,539,497 & 9.75\\
character[] & 4,576 & 3.82 & 88,233,523 & 3.37\\
class<`matrix`> & 4,522 & 3.77 & 32,452,555 & 1.24\\
class<`data.frame`> & 2,781 & 2.32 & 7,749,094 & 0.30\\
\bottomrule
\end{tabular}

% END Autogenerated
\label{tbl:corpus-type-breakdown}
\caption{Top types of contract assertion failures.}
\end{table}
